Supplementary Material

### 2022-11-12 paper/paper/SPIE/article.tex
[WIP]

The AUROC metric was chosen because it handles class imbalances present in the
various datasets, and is naturally normalized and interpretable such that
mAUROC \(<\) 0.5 indicates predictive performance worse than guessing, and
mAUROC \(>\) 0.5 indicates positive predictive
utility\cite{mandrekarReceiverOperatingCharacteristic2010}. However, we also
collected as performance metrics the accuracy, adjusted accuracy (accuracy
minus the proportion of the largest class), and the f1 score, and these metrics
are available with all the raw data for this study \ref{sec:data-and-code}.

the different

@@

Summarizing this quantity of data requires some caution. Measures of location or
scale, even when robust, are, for the most part, misleading and uninformative
when distributions are skewed. In most cases we find skewed mAUROC
distributions, and so have chosen primarily to present our findings visually,
with kernel density estimates. However, we also provide overall numerical
summaries of the feature performances in Table \ref{tab:numerical}.

\begin{table}[h!]
\caption{Numerical summaries of feature mAUROCs across predictable comparisons,
and all combinations of analytic choices, sorted by 95\% percentile (robust
max) value. Bold values indicate column "best" values, when reasonable. We
note, however, that aggregating across datasets in this manner can be deeply
misleading\cite{brabecModelEvaluationNonconstant2020}, and thus the values here
are intended solely as supplements to distribution plots. }
\label{tab:numerical}
\small
\centering

@@

In addition, each analytic factor could produce an interaction. For example, it
could be that trimming impacts the overall mAUROC distribution only at a
certain preprocessing level and for certain feature slicing. However, there are
too many potential interactions to present, so we limit our presentation to
main effects.  As there are a large number of features, we also summarize
patterns of results across two abstract groupings of similar features
("coarse", and "fine") shown in Table \ref{tab:features}.

It is also important to exclude a classification task or an analytic factor
from summaries if there is no evidence that the classification task is solvable
in general, or for that factor. We take as lack of evidence of solvability an
mAUROC distribution that is either (1) roughly symmetric and has mean and
median close to 0.5, or (2) with median and mode less than 0.5.

@@

\subsubsection{Non-Predictable Comparisons}
Based on the mAUROC distributions for either coarsely-grouped (Figure
\ref{fig:main-results}) or finely-grouped features (Figure
S\ref{S-fig:feature-group-all}), neither bilinguality nor depression were
predictable by either baseline or eigenfeatures. Additionally, the \code{pain v
duloxetine} comparison in the \textit{Osteo} dataset, and the
\code{trait\_attend v trait\_nonattend} conditions ("WeeklyAttend" in Figures
\ref{fig:main-results} and S\ref{S-fig:feature-group-all}) in the \textit{Attention}
dataset were also not meaningfully predictable by any feature. As such, we exclude
these classification tasks from further figures and discussion.

@@

@@

\subsection{Largest mAUROCs}
Considering the various analytic choices as tunable parameters, it makes sense
to examine the \textit{largest} portion of mAUROCs as an indication of the
maximum predictive potential of the eigenfeatures. In this case, it is clear
that eigenfeatures using RMT features almost always had the highest potential
predictive utility (Figure \ref{fig:largest}). Figure S\ref{S-fig:fine-largest}
shows that this was primarily due to either the "rmt only" or "rmt + eigs"
features (see Table \ref{tab:features}). However, these features were also most
likely to cause poor performance and overfitting (indicated by an mAUROC \(<\)
0.5; Figure S\ref{S-fig:fine-smallest}).

Examining these RMT features more closely (Figure S\ref{S-fig:fine-degree}),
it is clear these features are more predictive mostly due to the unfolding
procedure: the mAUROC distributions with largest medians and modes are due to
features that include the unfolded eigenvalues. Combined features that use
the unfolded eigenvalues plus some other RMT eigenfeature have identical
mAUROC distributions to those using the unfolded eigenvalues alone.

\subsection{Effect of Preprocessing}
At a coarse level, slice time correction followed by motion correction tended
to slightly increase the predictive utility of the eigenfeatures (Figure
\ref{fig:preproc}) relative to brain extraction only. Subsequent registration
following these steps did not generally impact mAUROC distributions further,
except in the \code{task\_attend v task\_nonattend} comparison, where
registration reduced the predictive utility of the eigenfeatures (Figure
\ref{fig:preproc}, second-last column). When examining features more finely, it
is clear that preprocessing most impacts the mAUROC distribution of the largest
and central eigenvalues ("eigs middle" and "eigs max" in Figure
S\ref{S-fig:fine-preproc}).

\subsection{Effect of Classifier}
mAUROC distributions were generally similar across coarse or fine feature
groupings (Figure S\ref{S-fig:coarse-classifier}, S\ref{S-fig:fine-classifier}).
In the \textit{Attention} \code{task\_attend v task\_nonattend} condition, eigenfeatures were modally predictive only with
the random forest classifier, whereas in the \textit{Aging} data, a SVC was least
likely to have mAUROC \(<\) 0.5. Overall, however, differences in the mAUROC distributions
due to classifier were small.

@@

\subsection{Effect of Slicing}
Features were slightly more predictive when using either the largest 5 or 10\% or middle
25\% ("max-05", "max-10" and "mid-25" in Figure
\ref{S-fig:best-params-slicing}), and least predictive when features primarily
involved the smallest eigenvalues.

@@

In this study, eigenfeatures inspired by RMT and derived from the eigenvalues
of the full, whole-brain voxelwise fMRI correlation matrix, were found to have
predictive utility across a wide variety of phenomena and analytic choices.
Compared to simple baseline reductions of the fMRI data, these eigenfeatures
had more consistent predictive utility, and a higher maximum predictive
potential (Figures \ref{fig:main-results}-\ref{fig:largest}). In addition to
evidence from previous studies
\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
this suggests RMT may be a useful analytic and theoretical tool for
understanding functional connectivity.

>>>>

The AUROC metric was chosen because it handles class imbalances present in the
various datasets, and is naturally normalized and interpretable such that
mAUROC \(<\) 0.5 indicates predictive performance worse than guessing, and
mAUROC \(>\) 0.5 indicates positive predictive
utility\cite{mandrekarReceiverOperatingCharacteristic2010}. However, we also
collected as performance metrics the accuracy, adjusted accuracy (accuracy
minus the proportion of the largest class), and the F1 score, and a complete
table with these additional metrics is available with the data for this study
(Section \ref{sec:data-and-code}).

@@

Summarizing this quantity of data requires some caution. Measures of location or
scale, even when robust, are, for the most part, misleading and uninformative
when distributions are skewed. In most cases we find skewed mAUROC
distributions, and so have chosen primarily to present our findings visually,
with kernel density estimates.

Because aggregating across datasets with different class imbalances can
be extremely misleading, even with a metric like the mAUROC\cite{brabecModelEvaluationNonconstant2020},
we restrict such summaries to Table \ref{tab:numerical} only, and note that
the table is intended as a \textit{supplement only} to the distribution plots: no conclusions
should be drawn about feature performance from the table values alone.

\begin{table}[h!]
\caption{Numerical summaries of feature mAUROCs across predictable comparisons,
and all combinations of analytic choices, sorted by 95\% percentile (robust
max) value. Bold values indicate column "best" values, when reasonable.}
\label{tab:numerical}
\small
\centering

@@

In addition, a large number of interactions are possible between each analytic
factor in this study. For example, it could be that trimming impacts the
overall mAUROC distribution only at a certain preprocessing level and for
certain feature slicing. However, as there are too many potential interactions
to present, we limit our presentation to main effects.

To aid in summarizing the performances of the large number of different
features, we also summarize patterns of results across two abstract groupings
of similar features ("coarse", and "fine") shown in Table \ref{tab:features}.
Thus, for example, figures depicting the coarse feature grouping "eigs" in fact
depict all observed mAUROC values for the the fine feature groupings of raw,
max, middle, and smoothed eigenvalues, and the fine feature grouping "eigs
smooth", for example, depicts all observed mAUROC values for both the
\code{eigs\_savgol} and \code{eigs\_smooth} features.

@@

\subsubsection{Non-Predictable Comparisons}

It is important, to reduce figure clutter and complexity, to exclude a
classification task or an analytic factor from summaries if there is no
evidence that the classification task is solvable in general, or for that
factor. We take as lack of evidence of solvability an mAUROC distribution that
is either (1) roughly symmetric and has mean and median close to 0.5, i.e.
performance appears random, or (2) with median and mode less than 0.5.

Based on the mAUROC distributions for either coarsely-grouped (Figure
\ref{fig:main-results}) or finely-grouped features (Figure
S\ref{S-fig:feature-group-all}), neither bilinguality nor depression were
predictable by either baseline or eigenfeatures. Additionally, the \code{pain v
duloxetine} comparison in the \textit{Osteo} dataset, and the
\code{trait\_attend v trait\_nonattend} conditions ("WeeklyAttend" in Figures
\ref{fig:main-results} and S\ref{S-fig:feature-group-all}) in the \textit{Attention}
dataset were also not meaningfully predictable by any feature. As such, we exclude
these classification tasks from further figures and discussion.

@@

\subsection{Largest mAUROCs}
Considering the various analytic choices as tunable parameters, it makes sense
to examine the \textit{largest} portion of mAUROCs as an indication of the
maximum predictive potential of the eigenfeatures. In this case, it is clear
that eigenfeatures using RMT features almost always had the highest potential
predictive utility (Figure \ref{fig:largest}). Figure S\ref{S-fig:fine-largest}
shows that this was primarily due to either the "rmt only" or "rmt + eigs"
features (see Table \ref{tab:features}). However, RMT eigenfeatures were also most
likely to cause poor performance and overfitting (indicated by an mAUROC \(<\)
0.5; Figure S\ref{S-fig:fine-smallest}).

Examining these RMT features more closely, it is clear these features'
performance distributions differ mostly due to the unfolding procedure. That
is, combined features that used the unfolded eigenvalues plus some other RMT
eigenfeature tended to have visually-indistinguishable mAUROC distributions to
those using the unfolded eigenvalues alone (Figures
S\ref{S-fig:fine-trim}-\ref{S-fig:fine-degree}). The mAUROC distributions of
these features differed most strongly in the tails (Figures
S\ref{S-fig:fine-largest}-\ref{S-fig:fine-smallest}).

\subsection{Effect of Preprocessing}
At a coarse level of feature grouping, slice time correction followed by motion
correction tended to slightly increase the predictive utility of the
eigenfeatures (Figure \ref{fig:preproc}) relative to brain extraction only.
Subsequent registration following these steps did not generally impact mAUROC
distributions further, except in the \code{task\_attend v task\_nonattend}
comparison, where registration reduced the predictive utility of the
eigenfeatures (Figure \ref{fig:preproc}, second-last column). When examining
features more finely, it is clear that preprocessing most impacts the mAUROC
distribution of the largest and central eigenvalues ("eigs middle" and "eigs
max" in Figure S\ref{S-fig:fine-preproc}).

@@

\subsection{Effect of Classifier}
Within any feature grouping (coarse or fine), and within a classification task,
mAUROC distributions were generally similar across classifiers (Figure
S\ref{S-fig:coarse-classifier}, S\ref{S-fig:fine-classifier}). Additionally,
these figures also show that within each classification task, choice of
classifier does not result in dramatic changes to the rough rank ordering of
features. E.g. if features are ranked on predictive utility, using the median,
modal, or bulk of the  mAUROC values, this rank ordering appears to remain
similar across classifiers.

In the \textit{Attention} \code{task\_attend v task\_nonattend} condition,
eigenfeatures were modally predictive only with the random forest classifier,
whereas in the \textit{Aging} data, a SVC was least likely to have mAUROC \(<\)
0.5. Overall, however, differences in the mAUROC distributions due to
classifier were small and inconsistent.

@@

\subsection{Effect of Slicing}
The best slice size and location depended complexly on the classification task
and feature, and few general summaries can be made of these interactions.
However, features including the full spectrum (e.g. raw eigenvalues, smoothed
eigenvalues, and their combinations with RMT eigenfeatures) were slightly more
predictive when using the largest portions ("max-XX", rows in Figure
S\ref{S-fig:best-params-slicing}), and usually least predictive when features
primarily involved the smallest or central portions.

@@

In this study, eigenfeatures inspired by RMT and derived from the eigenvalues
of the full, whole-brain voxelwise fMRI correlation matrix were found to have
predictive utility across a wide variety of phenomena and analytic choices.
Compared to simple baseline reductions of the fMRI data, these eigenfeatures
had more consistent predictive utility, and a higher maximum predictive
potential (Figures \ref{fig:main-results}-\ref{fig:largest}). In addition to
evidence from previous studies
\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
this suggests RMT may be a useful analytic and theoretical tool for
understanding functional connectivity.

However, mAUROC values observed in this study were highly sensitive to , and mean and
median mAUROC v

### 2022-11-12 paper/paper/SPIE/supplementary.tex
[WIP]

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{coarse_feature_group_by_predictive_subgroup_classifier.png}
\end{center}
\caption
{ \label{fig:coarse-classifier}
AUROC distributions across coarse feature groupings and predictable comparison tasks, by classifier.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_group_by_predictive_subgroup_classifier.png}
\end{center}
\caption
{ \label{fig:fine-classifier}
AUROC distributions across fine feature groupings and predictable comparison tasks, by classifier.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_largest_by_subgroup.png}
\end{center}
\caption
{ \label{fig:fine-largest}
Distributions of largest 500 mAUROCs across fine feature grouping, by comparison task.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_smallest_by_subgroup.png}
\end{center}
\caption
{ \label{fig:fine-smallest}
Distributions of smallest 500 mAUROCs across fine feature groupings, by comparison task.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_feature_auroc_by_trim.png}
\end{center}
\caption
{ \label{fig:fine-trim}
Distributions of mAUROCs for unfolding-dependent RMT features, by trimming.}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{best_rmt_params_by_subgroup.png}
\end{center}
\caption
{ \label{fig:best-params}
Distributions of mAUROCs when choosing ``best'' analytic choices and features.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{best_rmt_params_by_subgroup_and_slicing.png}
\end{center}
\caption
{ \label{fig:best-params-slicing}
Distributions of mAUROCs when choosing ``best'' analytic choices and features, by slicing.}
\end{figure}

>>>>

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_largest_by_subgroup.png}
\end{center}
\caption
{ \label{fig:fine-largest} Distributions of largest 500 mAUROCs across fine
feature grouping, by comparison task. Note "rmt only" and "rmt + eigs" features
tend to have the best possible performances across predictable tasks. }
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_smallest_by_subgroup.png}
\end{center}
\caption
{ \label{fig:fine-smallest} Distributions of smallest 500 mAUROCs across fine
feature groupings, by comparison task. Note "rmt only" and "rmt + eigs"
features tend to have the worse possible performances across predictable
tasks.}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{coarse_feature_group_by_predictive_subgroup_classifier.png}
\end{center}
\caption
{ \label{fig:coarse-classifier} AUROC distributions across coarse feature
groupings and predictable comparison tasks, by classifier. Note the general
similarity of each distribution within a particular classification task
(column) and within each feature grouping.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_group_by_predictive_subgroup_classifier.png}
\end{center}
\caption
{ \label{fig:fine-classifier} AUROC distributions across fine feature groupings
and predictable comparison tasks, by classifier. Note the general similarity of
each distribution within a particular classification task (column) and within
each feature grouping. Note also that, within a classification task (column),
that the rank ordering of features, based on wither the median, mode, or mean,
does not change dramatically or consistently from classifier to classifier. }
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_feature_auroc_by_trim.png}
\end{center}
\caption
{ \label{fig:fine-trim} Distributions of mAUROCs for unfolding-dependent RMT
features, by trimming. Note the tendency for a rightward shift in the mAUROC
distributions of the features involving the unfolded eigenvalues when using
largest or middle trimming (most dramatic in the Osteo nopain v duloxetine
condition). The impact of these trimming methods on the rigidity and level
variance features, however, was mixed (compare Vigilance data to Osteo nopain v
duloxetine condition.)}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_eigs_aurocs_by_subgroup_and_slicing.png}
\end{center}
\caption
{ \label{fig:best-params-slicing} Distributions of mAUROCs by slicing. Features
involving the full spectrum (raw eigenvalues, smoothed eigenvalues, and rmt +
eigs) sometimes have most positive mAUROC distributions when using the larger
eigenfeature values (first three columns) or middle values (Osteo nopain v pain
condition, Vigilance classification task).}
\end{figure}

% \begin{figure}[H]
% \begin{center}
% \includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{best_rmt_params_by_subgroup.png}
% \end{center}
% \caption
% { \label{fig:best-params}
% Distributions of mAUROCs when choosing ``best'' analytic choices and features.}
% \end{figure}
